Fig2: Candidate genes

### 2020-05-12 analysis/03_DESeq2part2.Rmd
attempting to save and read DEGs and vsd files

---
title: "03_datawrangle"
output: md_document
---

@@

head(candidatevsd)
```

>>>>

---
title: "03_DESeq2part2"
output: md_document
---

@@

head(candidatevsd)

## All DEGs (but currently only char DEGs)

```{r}
DEG_path <- "../results/DEseq2/treatment/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG)

# select only characterization plots
allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
allDEG$direction <- factor(allDEG$direction, levels = charlevels)

allDEG <- allDEG %>% drop_na()
```

### hypothesis testing

```{r}

DEG_path <- "../results/DEseq2/hypothesis/"   # path to the data	
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names	
DEG_pathfiles <- paste0(DEG_path, DEG_files)	
#DEG_files

allDEG2 <- DEG_pathfiles %>%	
  setNames(nm = .) %>% 	
  map_df(~read_csv(.x), .id = "file_name") %>% 	
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/hypothesis/'), "[", 2))  %>% 	
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 	
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%	
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%	
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%	
  mutate(comparison = paste(down,up, sep = "_")) %>%	
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%	
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%	
  dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 	
head(allDEG2)

allDEG2$tissue <- factor(allDEG2$tissue, levels = tissuelevel)

allDEG2$comparison <- factor(allDEG2$comparison, levels = c("eggs_chicks", "lo_hi"))	
allDEG2 <- allDEG2 %>% mutate(comparison = fct_recode(comparison, "lo vs. hi PRL   " = "lo_hi",	
                                                      "eggs vs. chicks" = "eggs_chicks"))	
allDEG2$direction <- factor(allDEG2$direction, levels = c("eggs", "chicks", "lo", "hi"))

PRLDEGs <- allDEG2 %>%	
  filter(tissue == "pituitary", comparison == "lo vs. hi PRL   ",	
         direction == "hi") %>%	
  arrange(desc(logpadj))

envDEGs <- allDEG2 %>%	
  filter(tissue == "pituitary", comparison == "eggs vs. chicks",	
         direction == "chicks") %>%	
  arrange(desc(logpadj))

hilogenes <- PRLDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)	
eggchickgenes <- envDEGs %>% head(20) %>% distinct(gene) %>% pull(gene)

hypothesisDEGs <- c(hilogenes, eggchickgenes) %>% unique()
```

## save files

```{r write}

write.csv(allDEG, "../results/03_hypothesisDEGs.csv")
write.csv(allDEG, "../results/03_allDEGs.csv")
write.csv(allvsd, "../results/03_allvsd.csv")
write.csv(candidatevsd, "../results/03_candidatevsd.csv")
```

### 2020-05-12 analysis/figure-1.Rmd
added code to print pdf next time
attempting to save and read DEGs and vsd files

@@

## DEGs

```{r allDEGs}
DEG_path <- "../results/DEseq2/treatment/"   # path to the data
DEG_files <- dir(DEG_path, pattern = "*DEGs") # get file names
DEG_pathfiles <- paste0(DEG_path, DEG_files)
#DEG_files

allDEG <- DEG_pathfiles %>%
  setNames(nm = .) %>% 
  map_df(~read_csv(.x), .id = "file_name") %>% 
  mutate(DEG = sapply(strsplit(as.character(file_name),'./results/DEseq2/treatment/'), "[", 2))  %>% 
  mutate(DEG = sapply(strsplit(as.character(DEG),'_diffexp.csv'), "[", 1))  %>% 
  mutate(tissue = sapply(strsplit(as.character(DEG),'\\.'), "[", 1)) %>%
  mutate(down = sapply(strsplit(as.character(DEG),'\\_'), "[", 3)) %>%
  mutate(up = sapply(strsplit(as.character(DEG),'\\_'), "[", 4)) %>%
  mutate(comparison = paste(down,up, sep = "_")) %>%
  mutate(sex = sapply(strsplit(as.character(sextissue),'\\_'), "[", 1)) %>%
  mutate(tissue = sapply(strsplit(as.character(sextissue),'\\_'), "[", 2)) %>%
dplyr::select(sex,tissue,comparison, direction, gene, lfc, padj, logpadj) 
head(allDEG)

# select only characterization plots
allDEG$tissue <- factor(allDEG$tissue , levels = tissuelevel)
allDEG$comparison <- factor(allDEG$comparison , levels = comparisonlevelschar)
allDEG$direction <- factor(allDEG$direction, levels = charlevels)

allDEG <- allDEG %>% drop_na()

@@

str(allDEG)
unique(allDEG$comparison)
```

@@

fig1 <- plot_grid(a,bc, d, nrow = 3, rel_heights = c(0.7,0.7,2),
                  labels = c("A"), label_size = 8)
fig1
```

>>>>

# uses count data from limma

@@

## Treatment specific DEGs

```{r}
allDEG <- read_csv("../results/03_allDEG.csv")
str(allDEG)
unique(allDEG$comparison)
```

## Sex and Tissue -related DEGs

```{r DEGs}

@@

```

@@

fig1 <- plot_grid(a,bc, d, nrow = 3, rel_heights = c(0.7,0.7,2),
                  labels = c("A"), label_size = 8)
fig1

pdf(file="../figures/fig1-1.pdf", width=7.25, height=7.25)
plot(fig1)
dev.off()
```

### 2020-05-12 analysis/figure-2.Rmd
added code to print pdf next time
attempting to save and read DEGs and vsd files

source("../R/themes.R")
source("../R/functions.R")
source("../R/wrangledata.R")

@@

```{r}
# load `candidatevsd` with `source("../R/wrangledata.R")`
head(candidatevsd)

@@

fig2 <- plot_grid(a,bcd, ncol = 1, labels = c("A"), label_size = 8, rel_heights = c(0.1,1))
fig2
```

>>>>

source("../R/themes.R")
source("../R/functions.R")

@@

```{r}
# load `candidatevsd` with `source("../R/wrangledata.R")`
candidatevsd <- read_csv("../results/03_candidatevsd.csv")
head(candidatevsd)

@@

fig2 <- plot_grid(a,bcd, ncol = 1, labels = c("A"), label_size = 8, rel_heights = c(0.1,1))
fig2

pdf(file="../figures/fig2-1.pdf", width=7.25, height=7.25)
plot(fig2)
dev.off()
```
